Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2019-12-06 manuscript/climate_sensitivity_draft.Rmd
working on LS comments

One fundamental question regarding forest responses to drought is what drives the observed tendency for large trees to suffer more during drought. @bennett_larger_2015 was the first to show greater growth reductions for larger trees on a global scale, and numerous subsequent studies have reinforced this finding (*e.g.*, @stovall_tree_2019; @hacket-pain_consistent_2016]. However, it has yet to be resolved which of several potential underlying mechanisms drive this pattern. First, taller trees have a greater biophysical challenge of lifting water greater distances against the effects of gravity and friction [@mcdowell_relationships_2011; @mcdowell_darcys_2015; @ryan_hydraulic_2006; @couvreur_water_2018], especially during drought [@zhang_size-dependent_2009]. Second, larger trees may have lower drought resistance because of their higher crown positions, where they are exposed to higher solar radiation, greater wind speeds, lower humidity, and lower CO_2_ concentrations [@koike_leaf_2001; @kunert_revised_2017 **REFS-KAT**). Subcanopy trees subsequently tend to fare better specifically due to the benefits of a buffered environment [@pretzsch_drought_2018]. Potentially counteracting the biophysical challenges faced by large trees, their larger root systems confer a potential advantage in terms of allowing greater access to water; however, it appears that this effect is usually insufficient to offset the costs of height and/or crown exposure. A final mechanism that could mediate tree size-related responses to drought is how hydraulic traits are distributed with respect to size [@meakem_role_2018; @liu_hydraulic_2019 [Roskilly et al. 2019](https://www.pnas.org/content/pnas/early/2019/06/14/1900734116.full.pdf)]. The pattern observed by @bennett_larger_2015 could be caused if the larger size classes were dominated by species less adapted to handle drought. Alternatively, larger size classes may be dominated by species that are inherently better adapted to greater biophysical challenges--as is the case in tropical moist forests of Panama, where larger size classes contain greater proportions of deciduous species [@condit_quantifying_2000; @meakem_role_2018]. Understanding the mechanisms driving the greater growth reductions of larger trees during drought will require sorting out the interactive effects of height, canopy position, soil water availability, and species' traits.

A second fundamental question regarding forest responses to drought is how species' traits -- alone and in interaction with tree size -- influence drought response [@elliott_forest_2015; @martinbenito_convergence_2015]. Separating species by ring porosity provides a first step toward identifying sensitivity, with diffuse-porous species tending to be more sensitive than ring-porous species [@kannenberg_linking_2019; @elliott_forest_2015; @friedrichs_species-specific_2009], but additional insight similar to @roskilly_conflicting_2019 is needed to better predict sensitivity at species or individual levels. Commonly-measured traits including wood density ($WD$) and leaf mass per area ($LMA$) have been linked to drought responses in temperate deciduous forests [@abrams_adaptations_1990; @guerfel_impacts_2009; @hoffmann_hydraulic_2011] and other forest biomes around the world [@greenwood_tree_2017]. However, the direction of response is not always consistent; wood density correlated negatively with drought tree performance in a broadleaf deciduous forest in the southeastern United States [@hoffmann_hydraulic_2011] but positively or with no effect at a global scale [@greenwood_tree_2017; @anderegg_hydraulic_2018). Hydraulic traits including water potentials at which percent loss conductivity surpass a certain threshold ($P50$, $P80$, $P88$) and hydraulic safety margin tend to be more successful at predicting drought performance [@anderegg_hydraulic_2018] but are time-consuming to measure and therefore infeasible for predicting or modeling drought responses in highly diverse forests (*e.g.*, in the tropics). More rapidly measurable leaf hydraulic traits with direct linkage to plant hydraulic function are emerging with potential to explain greater variation in plant distribution and function than the more commonly-measured $WD$ and $LMA$ [@medeiros_extensive_2019]. These include traits such as the fraction of conductivity lost to embolism, is another relevant metric (percent loss conductivity) [**REF??**], leaf area shrinkage upon dessication ($PLA_{dry}$; [@scoffoni_leaf_2014]), and the leaf water potential at turgor loss point ($\pi_{tlp}$), i.e., the water potential at which leaf wilting occurs [@bartlett_correlations_2016]. The abilities of both $PLA_{dry}$ and $\pi_{tlp}$ to explain tree performance under drought remains untested.

>>>>

One fundamental question regarding forest responses to drought is what drives the observed tendency for large trees to suffer more during drought. @bennett_larger_2015 was the first to show greater growth reductions for larger trees on a global scale, and numerous subsequent studies have reinforced this finding (*e.g.*, @stovall_tree_2019; @hacket-pain_consistent_2016]. However, it has yet to be resolved which of several potential underlying mechanisms drive this pattern. First, taller trees have a greater biophysical challenge of lifting water greater distances against the effects of gravity and friction [@mcdowell_relationships_2011; @mcdowell_darcys_2015; @ryan_hydraulic_2006; @couvreur_water_2018], especially during drought [@zhang_size-dependent_2009]. Second, larger trees may have lower drought resistance because of their higher crown positions, where they are exposed to higher solar radiation, greater wind speeds, lower humidity, and lower CO_2_ concentrations [@koike_leaf_2001; @kunert_revised_2017 **REFS-KAT**) **Lawren doesn't think CO2 would matter**. Subcanopy trees subsequently tend to fare better specifically due to the benefits of a buffered environment [@pretzsch_drought_2018]. Potentially counteracting the biophysical challenges faced by large trees, their larger root systems confer a potential advantage in terms of allowing greater access to water; however, it appears that this effect is usually insufficient to offset the costs of height and/or crown exposure. 
*A final mechanism that could mediate tree size-related responses to drought is how hydraulic traits are distributed with respect to size [@meakem_role_2018; @liu_hydraulic_2019 [Roskilly et al. 2019](https://www.pnas.org/content/pnas/early/2019/06/14/1900734116.full.pdf)]. **LS:—larger trees may have larger xylem conduits, which in turn may have less resistance to air seeding and subsequent embolism (Mark Olsen argument, right?) ** The pattern observed by @bennett_larger_2015 could be caused if the larger size classes were dominated by species less adapted to handle drought. Alternatively, larger size classes may be dominated by species that are inherently better adapted to greater biophysical challenges--as is the case in tropical moist forests of Panama, where larger size classes contain greater proportions of deciduous species [@condit_quantifying_2000; @meakem_role_2018]. Understanding the mechanisms driving the greater growth reductions of larger trees during drought will require sorting out the interactive effects of height, canopy position, soil water availability, and species' traits. *

Debates have also arisen regarding the traits that would influence tree growth responses to drought. Xylem architecture, for example, plays an important role, with diffuse porous species tending to be more drought-sensitive than ring-porous species [@kannenberg_linking_2019; @elliott_forest_2015; @friedrichs_species-specific_2009]. Commonly-measured traits including wood density ($WD$) and leaf mass per area ($LMA$) have been linked to drought responses in some temperate deciduous forests [@abrams_adaptations_1990; @guerfel_impacts_2009; @hoffmann_hydraulic_2011] and other forest biomes around the world [@greenwood_tree_2017], **but not in others (REFS)**. Indeed, the direction of response is not always consistent; wood density correlated negatively with drought tree performance in a broadleaf deciduous forest in the southeastern United States [@hoffmann_hydraulic_2011] but positively or with no effect at a global scale [@greenwood_tree_2017; @anderegg_hydraulic_2018] **and their role may be due to indirect correlations with other traits within life-history strategies (refs)**. Recent work has shown a great potential for hydraulic traits to predict growth and mortality responses. Hydraulic traits including water potentials at which percent loss conductivity surpass a certain threshold ($P50$, $P80$, $P88$) and hydraulic safety margin have enabled prediction of drought performance [@anderegg_hydraulic_2018] but are time-consuming to measure and therefore infeasible for predicting or modeling drought responses in highly diverse forests (*e.g.*, in the tropics). More rapidly measurable leaf hydraulic traits with direct linkage to plant hydraulic function are emerging with potential to explain greater variation in plant distribution and function than the more commonly-measured $WD$ and $LMA$ [@medeiros_extensive_2019]. These include leaf area shrinkage upon desiccation ($PLA_{dry}$; [@scoffoni_leaf_2014]) and the leaf water potential at turgor loss point ($\pi_{tlp}$), *i.e.,* the water potential at which leaf wilting occurs [@bartlett_correlations_2016]. The abilities of both $PLA_{dry}$ and $\pi_{tlp}$ to explain tree performance under drought remains untested.

**KAT is here with integrating LS comments**
